Report on a Safety Checker for Petri nets

### 2014-02-03 paper/workspace/main.tex
Added pseudocode for Invariant methods

\begin{center}
  \begin{tikzpicture}
    \node[state] (begin) {BEGIN};
    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, P)$};
    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
    \node[print, right=of satc] (noinv) {No Invariant};
    \node[action, below=of satc] (inv) {$A':=\text{Model}(C')$\\
      $I := \text{Inv}(N, A')$};
    \node[print, below=of inv] (printinv) {Invariant: $I$};
    \node[state, right=of noinv] (end1) {END};
    \node[state, below=of printinv] (end2) {END};

@@

    0        >= - b
    0        >= (- a) + b

@@

  Input:
    (S, T, E, M0) : Petri net
    P             : Property
  Output:
    C'            : Dual state constraints

Input:
 N = (S, T, E, M0)
   = ( (nc m cs1 cs2) (t1 t2 t3 t4) (e1 ... en) (2 1 0 0) )
 P = ( cs1 + cs2 <= 1 )

TODO: Unroll pseudocode?

* Subprocedure Inv:

  Input:
    (S, T, E, M0) : Petri net
    A             : Satisfying assignment for C'

@@

TODO: Unroll pseudocode?
\end{verbatim}

@@

\begin{verbatim}
* Property of minimization constraints C_M generated from A: 
    If A' satisfies C' u C_M, then Inv(N, A') uses less places than Inv(N, A)
\end{verbatim}

@@

    b_a + b_b + b_nbit2 < 3
    -------------------   |
                |         |
                |    number of places appearing in current invariant for A'
                |    = #{s | A'(s) > 0} = #{a, b, nbit2}
                |
                |
  number of places appearing in new invariant

@@

@@

\begin{verbatim}
* Subprocedure \text{MinConstraints}:

  Input:
    (S, T, E, M0) : Petri net
    A             : Assignment
  Output:
    C_M           : Minimization Constraints

Input:
 N   = (S, T, E, M0)
 A   : Satisfying assignment for C'

TODO: Unroll pseudocode?

@@

* Example:

>>>>

\begin{center}
  \begin{tikzpicture}
    \node[state] (begin) {BEGIN};
    \node[action, below=of begin] (c) {$C':=\mathcal C'(N, \neg P)$};
    \node[decision, below=of c] (satc) {$\text{SAT}(C')$};
    \node[print, right=of satc] (noinv) {No Invariant};
    \node[action, below=of satc] (inv) {$A':=\text{Model}(C')$\\
      $I := \text{Inv}(N, A')$};
    \node[print, below=of inv] (printinv) {Invariant: $I$};
    \node[state, right=of noinv] (end1) {END};
    \node[state, below=of printinv] (end2) {END};

@@

    0        >=    a  - b
    0        >= (- a) + b

@@

  Input:
    N = (S, T, E, M0)  : Petri net
    \neg P = (p_1,1 + ... + p_1,m_1 >= b_1 /\ p_2,1 + ... + p_2,m_2 >= b_2 /\
              ... /\ p_n,1 + ... + p_n,m_n >= b_n ) : Negated property
  Output:
    C'            : Dual state constraints

Input:
 N      = (S, T, E, M0)
        = ( (a b nbit2 ) (t1 t2) (e1 ... en) (1 0 0) )
\neg P  = ( a + b >= 2 )

Pseudocode:
\end{verbatim}

\begin{align*}
  C'(N, \neg P) =& \left( \bigwedge_{t \in T} \left( 0 \ge
                      \sum_{(t, s) \in E} s
                    - \sum_{(s, t) \in E} s \right) \right) \land
     \left( \sum_{s \in S} M_0(s) \cdot s <
       \sum_{i=1}^n b_i \cdot target_i \right) \land \\
     & \left( \bigwedge_{s \in S} \left ( s \ge 
       \sum_{i : s \in \{ p_{i,1}, \ldots, p_{i,m_i} \} } target_i \right) \right) \land
     \left( \bigwedge_{i=1}^n \left( target_i \ge 0 \right) \right)
\end{align*}

\begin{verbatim}
* Subprocedure Inv:

  Input:
    N = (S, T, E, M0) : Petri net
    A'                : Satisfying assignment for C'

@@

  Pseudocode:
\end{verbatim}

\begin{align*}
  I(N, A') =& \left( \sum_{s \in S} A'(s) \cdot s \le
                     \sum_{s \in S} A'(s) \cdot M_0(s) \right)
\end{align*}

@@

\begin{verbatim}
* Property of minimization constraints C_M generated from A': 
    If A'' satisfies C' u C_M, then Inv(N, A'') uses less places than Inv(N, A')
\end{verbatim}

@@

    b_a + b_b + b_nbit2 < 3
    -------------------   |
                |         |
                |    number of places appearing in current invariant for A'
                |    = #{s | A'(s) > 0} = #{a, b, nbit2}
                |
  number of places appearing in new invariant

@@

* Subprocedure \text{MinConstraints}:

  Input:
    N = (S, T, E, M0)  : Petri net
    A'                 : Satisfying assignment for C'
  Output:
    C_M                : Minimization Constraints

@@

\begin{align*}
  C_M(N, A') =& \left( \bigwedge_{s \in S} \left(
      (s > 0 \Rightarrow b_s = 1) \land (s = 0 \Rightarrow b_s = 0)
    \right) \right) \land
    \left( \sum_{s \in S} b_s < \sum_{s \in S : A'(s) > 0} 1 \right)
\end{align*}

@@

* Method Invariant by Refinement
\end{verbatim}

\begin{tikzpicture}
  \node[state] (begin) {BEGIN};
  \node[action, below=of begin] (c) {$C:=\mathcal C(N)$\\
                                        $D:=\{\}$};
  \node[decision, below=of c] (satc) {$\text{SAT}(C \cup \{\neg P\})$};
  \node[action, below=of satc] (modelc) {$A:=\text{Model}(C \cup \{\neg P\})$\\
                                        $C_{\theta}:=\text{TrapConditions}(N, A)$};
  \node[decision, below=of modelc] (satctheta) {$\text{SAT}(C_\theta)$};
  \node[action, below=of satctheta] (modelctheta) {$A_\theta:=\text{Model}(C_\theta)$\\
                                                $\delta:=\Delta(A_\theta)$\\
                                                $C:=C \cup \{\delta\}$\\
                                                $D:=D \cup \{\delta\}$};
  \node[print, right=of satctheta] (noinv2) {No Invariant};
  \node[state, below=of noinv2] (end3) {END};

  \node[action, right=of satc] (cprime) {$C':=\mathcal C'(N, \neg P \land D)$};
  \node[decision, right=of cprime] (satcprime) {$\text{SAT}(C')$};
  \node[action, below=of satcprime] (inv) {$A':=\text{Model}(C')$\\
                                        $I:=D \land \text{Inv}(N, A')$};
  \node[print, below=of inv] (printinv) {Invariant: $I$};
  \node[print, right=of satcprime] (noinv) {No Invariant};
  \node[state, below=of noinv] (end1) {END};
  \node[state, below=of printinv] (end2) {END};

  \draw (cprime) edge (satcprime);
  \draw (satcprime) edge node[above]{NO} (noinv);
  \draw (satcprime) edge node[right]{YES} (inv);
  \draw (inv) edge (printinv);
  \draw (noinv) edge (end1);
  \draw (printinv) edge (end2);

  \draw (begin) edge (c);
  \draw (c) edge coordinate[pos=.5] (edgein) (satc);
  \draw (satc) edge node[above]{NO} (cprime);
  \draw (satc) edge node[right]{YES} (modelc);
  \draw (modelc) edge (satctheta);
  \draw (satctheta) edge node[above]{NO} (noinv2);
  \draw (noinv2) edge (end3);
  \draw (satctheta) edge node[right]{YES} (modelctheta);
  \draw (modelctheta.south) -- ([yshift=-0.5cm] modelctheta.south)
  -| ([xshift=-2cm] modelctheta.west) |- (edgein);
\end{tikzpicture}

\begin{verbatim}
* Example:
